Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-03-11 supplement.tex

We performed an initial round of optimization including all six populations
with a family of models including single- and multiple-stem scenarios as
described in the previous section. We identified parameters that reached
consistent values across all models. These included the timings of recent or
non-central divergences and admixture events:
\begin{itemize}
    \item East/West African population split: 60ka
    \item East African/European split: 50 ka
    \item Neanderthal introgression to Eurasian branch: 45 ka
    \item Neanderthal/human split: 550 ka
    \item Eurasian back-to-African migration: 12 ka
\end{itemize}
When testing multiple variations of the more complex models, we kept these
values fixed (see Table~\ref{tab:fixed-params} for a complete list of all fixed
parameters). This reduced the potential for overfitting the more complex
models, while reducing the computational cost of optimization.

Our models also included recent events to account for known migrations,
admixtures, and growths and declines in effective population sizes. Many of
these parameters were fixed based on previous historical, genetic, or
anthropological research, namely
\begin{itemize}
    \item East African pastoralist to South African admixture: 2 ka \cite{}
    \item European to South African admixture: 10 generations ago \cite{}
    \item Mende population expansion and Gumuz population decline: 5 ka \cite{}
    \item South African population decline following colonial admixture: 9 generations ago \cite{}
\end{itemize}
\comment{This could be organized with a bit more structure. E.g. by region, or by timepoint}
While the dates of these events were fixed, the sizes and proportions were
allowed to vary in the fits. The total number of parameters that were
ultimately inferred were 21-31, depending on the complexity
of the model.

@@

\begin{table}[ht]
\caption{
    \label{tab:supp-single-origin}
    \textbf{Best-fit parameters from the single-origin model.}
    This model includes a single stem population that splits into contemporary human
    groups. Fixed parameters are specifed in Table~\ref{tab:fixed-params}.
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -189,434.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & Std. err.\\
    \midrule
    $N_e$ & Ancestral effective population size & 10198 & 403 \\
    $N_{MH}$ & Size of human lineage between Neanderthal and Nama splits & 21111 & 529 \\
    $N_{Nama_0}$ & Initial Nama size & 10224 & 370 \\
    $N_{Nama_F}$ & Final Nama size & 222 & 9 \\
    $N_{MSL_0}$ & Initial Mende size & 17211 & 769 \\
    $N_{MSL_F}$ & Final Mende size & 16822 & 606 \\
    $N_{EA}$ & Size of East African branch & 7139 & 273 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3831 & 131 \\
    $N_{EP}$ & East African agriculturist size & 13033 & 491 \\
    $N_{GBR_0}$ & Initial British size & 846 & 33 \\
    $N_{GBR_F}$ & Final British size & 12121 & 507 \\
    $N_{Neand}$ & Neanderthal size & 1867 & 105 \\
    $T_{Nama}$ & Nama split time (years) & 110400 & 2525 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $2.82\times10^{-5}$ & $0.158\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.94\times10^{-5}$ & $0.197\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $18.76\times10^{-5}$ & $0.764\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $4.42\times10^{-5}$ & $0.239\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $41.28\times10^{-5}$ & $1.33\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12 ka ($1-f$ from Gumuz) & 0.658 & 0.0039 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2 ka & 0.279 & 0.0039 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.150 & 0.0019 \\
    \bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{
    \label{tab:supp-continuous-migration-no-mig}
    \textbf{Best-fit parameters from the Continuous-Migration model without stem migration.}
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -126,644. This model allows for
    continuous migration between Stem 2 and contemporary populations, but disallows
    migration between Stem 2 and Stem 1, which later splits into branches leading
    to South, East, and West African populations.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & Std. err.\\
    \midrule
    $N_e$ & Ancestral effective population size & 10543 & 460 \\
    $N_{stem1}$ & Size of stem 1 lineage between Neanderthal and Nama splits & 15297 & 1227 \\
    $N_{stem2}$ & Size of stem 2 lineage & 9156 & 573 \\
    $N_{Nama_0}$ & Initial Nama size & 11329 & 423 \\
    $N_{Nama_F}$ & Final Nama size & 219 & 11 \\
    $N_{MSL_0}$ & Initial Mende size & 11187 & 644 \\
    $N_{MSL_F}$ & Final Mende size & 24056 & 945 \\
    $N_{EA}$ & Size of East African branch & 7185 & 293 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3746 & 184 \\
    $N_{EP}$ & East African agriculturist size & 13187 & 645 \\
    $N_{GBR_0}$ & Initial British size & 937 & 41 \\
    $N_{GBR_F}$ & Final British size & 11478 & 420 \\
    $N_{Neand}$ & Neanderthal size & 2285 & 113 \\
    $T_{stems}$ & Stem split time (years) & 459720 & 21704 \\
    $T_{Nama}$ & Nama split time (years) & 139725 & 3250 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $1.93\times10^{-5}$ & $0.360\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.43\times10^{-5}$ & $0.309\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $20.8\times10^{-5}$ & $0.97\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $4.46\times10^{-5}$ & $0.171\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $35.4\times10^{-5}$ & $1.85\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12 ka ($1-f$ from Gumuz) & 0.649 & 0.0038 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2 ka & 0.268 & 0.0048 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.153 & 0.0022 \\
    $m_{stems}$ & Stem 1--stem 2 migration rate & $0$ & fixed \\
    $m_{stem2-Nama}$ & Stem 2--Nama migration rate & $4.33\times10^{-5}$ & $0.891\times10^{-5}$ \\
    $m_{stem2-MSL}$ & Stem 2--Mende migration rate & $11.3\times10^{-5}$ & $1.83\times10^{-5}$ \\
    $m_{stem2-EA}$ & Stem 2--East Africa migration rate & $2.77\times10^{-5}$ & $0.699\times10^{-5}$ \\
    \bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{
    \label{tab:supp-continuous-migration}
    \textbf{Best-fit parameters from the Continuous-Migration model.}
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -115,500.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & Std. err.\\
    \midrule
    $N_e$ & Ancestral effective population size & 7270 & 1777 \\
    $N_{stem1}$ & Size of stem 1 lineage between Neanderthal and Nama splits & 8256 & 1612 \\
    $N_{stem2}$ & Size of stem 2 lineage & 13547 & 2488 \\
    $N_{Nama_0}$ & Initial Nama size & 11939 & 2989 \\
    $N_{Nama_F}$ & Final Nama size & 221 & 54 \\
    $N_{MSL_0}$ & Initial Mende size & 9738 & 2479 \\
    $N_{MSL_F}$ & Final Mende size & 28150 & 6628 \\
    $N_{EA}$ & Size of East African branch & 7489 & 1841 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3728 & 915 \\
    $N_{EP}$ & East African agriculturist size & 13072 & 3246 \\
    $N_{GBR_0}$ & Initial British size & 959 & 231 \\
    $N_{GBR_F}$ & Final British size & 11822 & 2889 \\
    $N_{Neand}$ & Neanderthal size & 2670 & 591 \\
    $T_{stems}$ & Stem split time (years) & 1163072 & 390803 \\
    $T_{Nama}$ & Nama split time (years) & 134745 & 17775 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $0.98\times10^{-5}$ & $0.366\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.08\times10^{-5}$ & $1.02\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $21.4\times10^{-5}$ & $5.32\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $4.17\times10^{-5}$ & $1.02\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $33.6\times10^{-5}$ & $8.35\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12 ka ($1-f$ from Gumuz) & 0.642 & 0.0037 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2 ka & 0.255 & 0.0043 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.156 & 0.0021 \\
    $m_{stems}$ & Stem 1--stem 2 migration rate & $6.43\times10^{-5}$ & $1.05\times10^{-5}$ \\
    $m_{stem2-Nama}$ & Stem 2--Nama migration rate & $5.82\times10^{-5}$ & $1.60\times10^{-5}$ \\
    $m_{stem2-MSL}$ & Stem 2--Mende migration rate & $16.4\times10^{-5}$ & $4.19\times10^{-5}$ \\
    $m_{stem2-EA}$ & Stem 2--East Africa migration rate & $3.10\times10^{-5}$ & $0.901\times10^{-5}$ \\
    \bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{
    \label{tab:supp-merger-without-stem-migration}
    \textbf{Best-fit parameters from the Merger-Without-Stem-Migration model.}
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -107,652.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & Std. err.\\
    \midrule
    $N_e$ & Ancestral effective population size & 11258 & 326 \\
    $N_{stem1}$ & Size of stem 1 lineage between stem 1--stem 2 split and stem 1E--stem 1S split & 113 & 76 \\
    $N_{stem2}$ & Size of stem 2 lineage & 23984 & 1149 \\
    $N_{Nama_0}$ & Initial Nama and stem 1S size & 13134 & 384 \\
    $N_{Nama_F}$ & Final Nama size & 225 & 7.3 \\
    $N_{MSL_0}$ & Initial Mende size & 11856 & 322 \\
    $N_{MSL_F}$ & Final Mende size & 25558 & 987 \\
    $N_{EA}$ & Size of East African and stem 1E branch & 9136 & 246 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3385 & 102 \\
    $N_{EP}$ & East African agriculturist size & 13650 & 408 \\
    $N_{GBR_0}$ & Initial British size & 931 & 29 \\
    $N_{GBR_F}$ & Final British size & 12064 & 334 \\
    $N_{Neand}$ & Neanderthal size & 1935 & 91 \\
    $T_{stems}$ & Stem split time (years) & 420881 & 27380 \\
    $T_{stem1}$ & Stem 1 split time into stem 1E and stem 1S (years) & 367434 & 19952 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $0.361\times10^{-5}$ & $0.113\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.00\times10^{-5}$ & $0.130\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $19.5\times10^{-5}$ & $0.548\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $3.77\times10^{-5}$ & $0.152\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $37.1\times10^{-5}$ & $1.26\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12 ka ($1-f$ from Gumuz) & 0.647 & 0.0037 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2 ka & 0.257 & 0.0042 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.156 & 0.0021 \\
    $m_{stems}$ & Stem 1--stem 2 migration rate & $0$ & fixed \\
    $T_{Nama}$ & Time of Nama merger event & 117392 & 8253 \\
    $f_{stem 2 \rightarrow Nama}$ & Proportion of stem 2 ancsestry making up initial Nama lineage ($1-f$ from stem 1S) & 0.707 & 0.0086 \\
    $T_{EA}$ & Time of East Africa merger event & 94892 & 3648 \\
    $f_{stem 2 \rightarrow EA}$ & Proportion of stem 2 ancestry making up initial East Africa lineage ($1-f$ from stem 1E) & 0.481 & 0.0074 \\
    $T_{MSL}$ & Time of secondary admixture from stem 2 to Mende & 23922 & 570 \\
    $f_{stem 2 \rightarrow MSL}$ & Proportion of ancestry from secondary stem 2 admixture to Mende & 0.168 & 0.0036 \\
    \bottomrule
\end{tabular}
\end{table}

>>>>

We performed an initial round of optimization including all six populations
with a family of models including single- and multiple-stem scenarios as
described in the previous section. We identified parameters that reached
consistent values across all models. These included the timings of recent or
non-central divergences and admixture events:
\begin{itemize}
    \item East/West African population split: 60ka
    \item East African/European split: 50ka
    \item Neanderthal introgression to Eurasian branch: 45ka
    \item Neanderthal/human split: 550ka
    \item Eurasian back-to-African migration: 12ka
\end{itemize}
When testing multiple variations of the more complex models, we kept these
values fixed (see Table~\ref{tab:fixed-params} for a complete list of all fixed
parameters). This reduced the potential for overfitting the more complex
models, while reducing the computational cost of optimization.

Our models also included recent events to account for known migrations,
admixtures, and growths and declines in effective population sizes. Many of
these parameters were fixed based on previous historical, genetic, or
anthropological research, namely
\begin{itemize}
    \item East African pastoralist to South African admixture: 2ka \cite{}
    \item European to South African admixture: 10 generations ago \cite{}
    \item Mende population expansion and Gumuz population decline: 5ka \cite{}
    \item South African population decline following colonial admixture: 9 generations ago \cite{}
\end{itemize}
\comment{This could be organized with a bit more structure. E.g. by region, or by timepoint}
While the dates of these events were fixed, the sizes and proportions were
allowed to vary in the fits. The total number of parameters that were
ultimately inferred were 21-31, depending on the complexity
of the model.

@@

\begin{table}[ht]
\caption{
    \label{tab:supp-single-origin}
    \textbf{Best-fit parameters from the single-origin model.}
    This model includes a single stem population that splits into contemporary human
    groups. Fixed parameters are specifed in Table~\ref{tab:fixed-params}.
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -189,434.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & Std. err.\\
    \midrule
    $N_e$ & Ancestral effective population size & 10198 & 403 \\
    $N_{MH}$ & Size of human lineage between Neanderthal and Nama splits & 21111 & 529 \\
    $N_{Nama_0}$ & Initial Nama size & 10224 & 370 \\
    $N_{Nama_F}$ & Final Nama size & 222 & 9 \\
    $N_{MSL_0}$ & Initial Mende size & 17211 & 769 \\
    $N_{MSL_F}$ & Final Mende size & 16822 & 606 \\
    $N_{EA}$ & Size of East African branch & 7139 & 273 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3831 & 131 \\
    $N_{EP}$ & East African agriculturist size & 13033 & 491 \\
    $N_{GBR_0}$ & Initial British size & 846 & 33 \\
    $N_{GBR_F}$ & Final British size & 12121 & 507 \\
    $N_{Neand}$ & Neanderthal size & 1867 & 105 \\
    $T_{Nama}$ & Nama split time (years) & 110400 & 2525 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $2.82\times10^{-5}$ & $0.158\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.94\times10^{-5}$ & $0.197\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $18.76\times10^{-5}$ & $0.764\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $4.42\times10^{-5}$ & $0.239\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $41.28\times10^{-5}$ & $1.33\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12ka ($1-f$ from Gumuz) & 0.658 & 0.0039 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2ka & 0.279 & 0.0039 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.150 & 0.0019 \\
    \bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{
    \label{tab:supp-continuous-migration-no-mig}
    \textbf{Best-fit parameters from the Continuous-Migration model without stem migration.}
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -126,644. This model allows for
    continuous migration between Stem 2 and contemporary populations, but disallows
    migration between Stem 2 and Stem 1, which later splits into branches leading
    to South, East, and West African populations.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & Std. err.\\
    \midrule
    $N_e$ & Ancestral effective population size & 10543 & 460 \\
    $N_{stem1}$ & Size of stem 1 lineage between Neanderthal and Nama splits & 15297 & 1227 \\
    $N_{stem2}$ & Size of stem 2 lineage & 9156 & 573 \\
    $N_{Nama_0}$ & Initial Nama size & 11329 & 423 \\
    $N_{Nama_F}$ & Final Nama size & 219 & 11 \\
    $N_{MSL_0}$ & Initial Mende size & 11187 & 644 \\
    $N_{MSL_F}$ & Final Mende size & 24056 & 945 \\
    $N_{EA}$ & Size of East African branch & 7185 & 293 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3746 & 184 \\
    $N_{EP}$ & East African agriculturist size & 13187 & 645 \\
    $N_{GBR_0}$ & Initial British size & 937 & 41 \\
    $N_{GBR_F}$ & Final British size & 11478 & 420 \\
    $N_{Neand}$ & Neanderthal size & 2285 & 113 \\
    $T_{stems}$ & Stem split time (years) & 459720 & 21704 \\
    $T_{Nama}$ & Nama split time (years) & 139725 & 3250 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $1.93\times10^{-5}$ & $0.360\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.43\times10^{-5}$ & $0.309\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $20.8\times10^{-5}$ & $0.97\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $4.46\times10^{-5}$ & $0.171\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $35.4\times10^{-5}$ & $1.85\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12ka ($1-f$ from Gumuz) & 0.649 & 0.0038 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2ka & 0.268 & 0.0048 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.153 & 0.0022 \\
    $m_{stems}$ & Stem 1--stem 2 migration rate & $0$ & fixed \\
    $m_{stem2-Nama}$ & Stem 2--Nama migration rate & $4.33\times10^{-5}$ & $0.891\times10^{-5}$ \\
    $m_{stem2-MSL}$ & Stem 2--Mende migration rate & $11.3\times10^{-5}$ & $1.83\times10^{-5}$ \\
    $m_{stem2-EA}$ & Stem 2--East Africa migration rate & $2.77\times10^{-5}$ & $0.699\times10^{-5}$ \\
    \bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{
    \label{tab:supp-continuous-migration}
    \textbf{Best-fit parameters from the Continuous-Migration model.}
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -115,500.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & Std. err.\\
    \midrule
    $N_e$ & Ancestral effective population size & 7270 & 1777 \\
    $N_{stem1}$ & Size of stem 1 lineage between Neanderthal and Nama splits & 8256 & 1612 \\
    $N_{stem2}$ & Size of stem 2 lineage & 13547 & 2488 \\
    $N_{Nama_0}$ & Initial Nama size & 11939 & 2989 \\
    $N_{Nama_F}$ & Final Nama size & 221 & 54 \\
    $N_{MSL_0}$ & Initial Mende size & 9738 & 2479 \\
    $N_{MSL_F}$ & Final Mende size & 28150 & 6628 \\
    $N_{EA}$ & Size of East African branch & 7489 & 1841 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3728 & 915 \\
    $N_{EP}$ & East African agriculturist size & 13072 & 3246 \\
    $N_{GBR_0}$ & Initial British size & 959 & 231 \\
    $N_{GBR_F}$ & Final British size & 11822 & 2889 \\
    $N_{Neand}$ & Neanderthal size & 2670 & 591 \\
    $T_{stems}$ & Stem split time (years) & 1163072 & 390803 \\
    $T_{Nama}$ & Nama split time (years) & 134745 & 17775 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $0.98\times10^{-5}$ & $0.366\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.08\times10^{-5}$ & $1.02\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $21.4\times10^{-5}$ & $5.32\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $4.17\times10^{-5}$ & $1.02\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $33.6\times10^{-5}$ & $8.35\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12ka ($1-f$ from Gumuz) & 0.642 & 0.0037 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2ka & 0.255 & 0.0043 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.156 & 0.0021 \\
    $m_{stems}$ & Stem 1--stem 2 migration rate & $6.43\times10^{-5}$ & $1.05\times10^{-5}$ \\
    $m_{stem2-Nama}$ & Stem 2--Nama migration rate & $5.82\times10^{-5}$ & $1.60\times10^{-5}$ \\
    $m_{stem2-MSL}$ & Stem 2--Mende migration rate & $16.4\times10^{-5}$ & $4.19\times10^{-5}$ \\
    $m_{stem2-EA}$ & Stem 2--East Africa migration rate & $3.10\times10^{-5}$ & $0.901\times10^{-5}$ \\
    \bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{
    \label{tab:supp-merger-without-stem-migration}
    \textbf{Best-fit parameters from the Merger-Without-Stem-Migration model.}
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -107,652.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & Std. err.\\
    \midrule
    $N_e$ & Ancestral effective population size & 11258 & 326 \\
    $N_{stem1}$ & Size of stem 1 lineage between stem 1--stem 2 split and stem 1E--stem 1S split & 113 & 76 \\
    $N_{stem2}$ & Size of stem 2 lineage & 23984 & 1149 \\
    $N_{Nama_0}$ & Initial Nama and stem 1S size & 13134 & 384 \\
    $N_{Nama_F}$ & Final Nama size & 225 & 7.3 \\
    $N_{MSL_0}$ & Initial Mende size & 11856 & 322 \\
    $N_{MSL_F}$ & Final Mende size & 25558 & 987 \\
    $N_{EA}$ & Size of East African and stem 1E branch & 9136 & 246 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3385 & 102 \\
    $N_{EP}$ & East African agriculturist size & 13650 & 408 \\
    $N_{GBR_0}$ & Initial British size & 931 & 29 \\
    $N_{GBR_F}$ & Final British size & 12064 & 334 \\
    $N_{Neand}$ & Neanderthal size & 1935 & 91 \\
    $T_{stems}$ & Stem split time (years) & 420881 & 27380 \\
    $T_{stem1}$ & Stem 1 split time into stem 1E and stem 1S (years) & 367434 & 19952 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $0.361\times10^{-5}$ & $0.113\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.00\times10^{-5}$ & $0.130\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $19.5\times10^{-5}$ & $0.548\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $3.77\times10^{-5}$ & $0.152\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $37.1\times10^{-5}$ & $1.26\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12ka ($1-f$ from Gumuz) & 0.647 & 0.0037 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2ka & 0.257 & 0.0042 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.156 & 0.0021 \\
    $m_{stems}$ & Stem 1--stem 2 migration rate & $0$ & fixed \\
    $T_{Nama}$ & Time of Nama merger event & 117392 & 8253 \\
    $f_{stem 2 \rightarrow Nama}$ & Proportion of stem 2 ancsestry making up initial Nama lineage ($1-f$ from stem 1S) & 0.707 & 0.0086 \\
    $T_{EA}$ & Time of East Africa merger event & 94892 & 3648 \\
    $f_{stem 2 \rightarrow EA}$ & Proportion of stem 2 ancestry making up initial East Africa lineage ($1-f$ from stem 1E) & 0.481 & 0.0074 \\
    $T_{MSL}$ & Time of secondary admixture from stem 2 to Mende & 23922 & 570 \\
    $f_{stem 2 \rightarrow MSL}$ & Proportion of ancestry from secondary stem 2 admixture to Mende & 0.168 & 0.0036 \\
    \bottomrule
\end{tabular}
\end{table}

\begin{table}[ht]
\caption{
    \label{tab:supp-merger-with-stem-migration}
    \textbf{Best-fit parameters from the Merger-With-Stem-Migration model.}
    Inferred values are scaled to physical units assuming a generation time of
    29 years. This model gave a log-likelihood of -102,633.
}
\centering
\begin{tabular}[t]{rP{8cm}cc}
    \toprule
    Parameter & Description & Value & Std. err.\\
    \midrule
    $N_e$ & Ancestral effective population size & 11479 & 1369 \\
    $N_{stem1}$ & Size of stem 1 lineage between Neanderthal split and stem 1E--stem 1S split & 117 & 838 \\
    $N_{stem2}$ & Size of stem 2 lineage & 24393 & 6668 \\
    $N_{Nama_0}$ & Initial Nama size & 13211 & 1514 \\
    $N_{Nama_F}$ & Final Nama size & 223 & 31 \\
    $N_{MSL_0}$ & Initial Mende size & 11444 & 1165 \\
    $N_{MSL_F}$ & Final Mende size & 27417 & 4332 \\
    $N_{EA}$ & Size of East African Branch & 9077 & 1628 \\
    $N_{Gumuz_F}$ & Final Gumuz size & 3402 & 337 \\
    $N_{EP}$ & East African agriculturist size & 13506 & 1684 \\
    $N_{GBR_0}$ & Initial British size & 953 & 122 \\
    $N_{GBR_F}$ & Final British size & 12406 & 1678 \\
    $N_{Neand}$ & Neanderthal size & 2416 & 235 \\
    $T_{stems}$ & Stem split time (years) & 1442022 & 426449 \\
    $T_{stem1}$ & Stem 1S--stem 1E split time (years) & 479401 & 166339 \\
    $m_{Nama-MSL}$ & Nama--Mende symmetric migration rate & $0.712\times10^{-5}$ & $0.401\times10^{-5}$ \\
    $m_{Nama-EA}$ & Nama--East Africa symmetric migration rate & $4.35\times10^{-5}$ & $0.912\times10^{-5}$ \\
    $m_{MSL-EA}$ & Mende--East Africa migration rate & $19.8\times10^{-5}$ & $2.57\times10^{-5}$ \\
    $m_{EA-GBR}$ & East Africa--Europe migration rate & $3.87\times10^{-5}$ & $0.550\times10^{-5}$ \\
    $m_{EA-EA}$ & Intra-East Africa migration rate & $35.9\times10^{-5}$ & $5.36\times10^{-5}$ \\
    $f_{GBR \rightarrow EP}$ & Ancestry proportion of East African agriculturalists from GBR 12ka ($1-f$ from Gumuz) & 0.640 & 0.0075 \\
    $f_{EP \rightarrow Nama}$ & Ancestry proportion from EA pastoralists to Nama 2ka & 0.257 & 0.0049 \\
    $f_{GBR \rightarrow Nama}$ & Ancestry proportion from Europeans to Nama 10 generations ago & 0.157 & 0.0031 \\
    $m_{stems}$ & Stem 1--stem 2 migration rate & $11.6\times10^{-5}$ & $8.74\times10^{-5}$ \\
    $T_{Nama}$ & Time of Nama merger event & 118547 & 28170 \\
    $f_{stem 2 \rightarrow Nama}$ & Proportion of stem 2 ancsestry making up initial Nama lineage ($1-f$ from stem 1S) & 0.714 & 0.067 \\
    $T_{EA}$ & Time of East Africa merger event & 98083 & 8865 \\
    $f_{stem 2 \rightarrow EA}$ & Proportion of stem 2 ancestry making up initial East Africa lineage ($1-f$ from stem 1E) & 0.495 & 0.059 \\
    $T_{MSL}$ & Time of secondary admixture from stem 2 to Mende & 25119 & 641 \\
    $f_{stem 2 \rightarrow MSL}$ & Proportion of ancestry from secondary stem 2 admixture to Mende & 0.181 & 0.0085 \\
    \bottomrule
\end{tabular}
\end{table}
